---
title: "Food Access in different Regions"
subtitle: "Report"
format: html
editor: visual
execute:
  echo: false
---

```{r}
#| label: read libraries

library(readr)
library(dplyr)
library(ggplot2)
library(tidyverse)
library(tidymodels)


food_access <- read_csv("data/food_access.csv")
```

```{r}
#| label: creating proportions and mutating data

Midwest <- c("Ohio", "Indiana", "Michigan", "Illinois", "Missouri", "Wisconsin", "Minnesota", "Iowa", "Kansas", "Nebraska", "South Dakota", "North Dakota")

Northeast <- c ("Maine", "Massachusetts", "Rhode Island", "Connecticut", "New Hampshire", "Vermont", "New York", "Pennsylvania", "New Jersey", "Delaware", "Maryland", "District of Columbia")

Southwest <- c ("Texas", "Oklahoma", "New Mexico", "Arizona")

Southeast <- c ("West Virginia", "Virginia", "Kentucky", "Tennessee", "North Carolina", "South Carolina", "Georgia", "Alabama", "Mississippi", "Arkansas", "Louisiana", "Florida")

West <- c("Colorado", "Wyoming", "Montana", "Idaho", "Washington", "Oregon", "Utah", "Nevada", "California", "Alaska", "Hawaii")

food_access_reg <- food_access |>
  mutate (region = case_when(State %in% Northeast ~ "Northeast", State %in% Southeast ~ "Southeast", State %in% Midwest ~ "Midwest", State %in% Southwest ~ "Southwest", State %in% West ~ "West")) 


food_access_reg <-food_access_reg |>
  mutate (proportion = `Low Access Numbers.Low Income People.10 Miles` / `Housing Data.Total Housing Units`, proportion_vechicle = `Low Access Numbers.Low Income People.10 Miles`, proportion_vehicle = `Vehicle Access.10 Miles`/`Housing Data.Total Housing Units`, proportion_seniors = `Low Access Numbers.Seniors.10 Miles`/`Housing Data.Total Housing Units`, proportion_group = `Housing Data.Residing in Group Quarters`/`Housing Data.Total Housing Units`)
```


```{r}
#| label: food access dataset

food_access<- food_access |>
  mutate(percentage = `Vehicle Access.1 Mile`/ `Housing Data.Total Housing Units`) 

food_access |>
  group_by(State) |>
  summarise(mean = mean(percentage))

```

```{r}
#| label: creating new column

# Read in the data

# Rename the Housing Data.Residing in Group Quarters column
food_access <- food_access %>% rename(group_quarters = `Housing Data.Residing in Group Quarters`)

# Create a scatterplot to visualize the relationship between the two variables
ggplot(data = food_access, aes(x = group_quarters, y = `Vehicle Access.1 Mile`)) +
  geom_point() +
  labs(x = "Percentage of Population Residing in Group Quarters", y = "Number of Housing Units without Vehicle Access Beyond 1 Mile from a Supermarket") +
  ggtitle("Relationship between Group Quarters and Housing Units without Vehicle Access") 

# Perform a linear regression to quantify the relationship

  
model1 <- linear_reg() |>
  set_engine("lm") |>
  fit(`Vehicle Access.1 Mile` ~ group_quarters , data = food_access)

model1 |>
  glance() |>
  pull(adj.r.squared)

```

the adjusted R-squared value of 0.43 suggests that the predictor variable "Vehicle Access.1 Mile" accounts for approximately 43% of the variability in the response variable "group_quarters", after adjusting for the effect of other predictor variables in the model.

This means that there is a moderate positive linear relationship between the percentage of housing units without vehicle access beyond 1 mile from a supermarket (as represented by the predictor variable "Vehicle Access.1 Mile") and the percentage of tract population residing in group quarters (as represented by the response variable "group_quarters").



```{r}
#| label: x

summary(food_access[,c("Population", "group_quarters", "Housing Data.Total Housing Units", "Vehicle Access.1 Mile")])

# Check for missing values
sum(is.na(food_access))

ggplot(food_access, aes(x= group_quarters)) +
  geom_histogram(binwidth=10, fill="blue", color="white") +
  labs(title="Distribution of population residing in group quarters", x="Population in group quarters", y="Frequency")

ggplot(food_access, aes(x=`Vehicle Access.1 Mile`, y= group_quarters)) +
  geom_point() +
  labs(title="Scatterplot of vehicle access beyond 1 mile against population in group quarters", x="Vehicle Access.1 Mile", y="Population in group quarters")
```

```{r}
#| label: summarizing mean

food_access |>
  group_by(State) |>
  summarise(total = mean(group_quarters)) 


```

district of columbia has the most living in group quarters

```{r}
#| label: linear representation

food_access |>
  filter(State == "District of Columbia") |>
  ggplot(aes(x = group_quarters, y = `Vehicle Access.1 Mile`)) +
  geom_point() +
  geom_smooth(method = "lm" , se = F) 

```

```{r}
#| label: linear regression

food_access_DC <- food_access |>
  filter(State == "District of Columbia")

linear_reg() |>
  set_engine("lm") |>
  fit(`Low Access Numbers.Low Income People.1 Mile` ~ `Vehicle Access.1 Mile` + `Low Access Numbers.Seniors.1 Mile`, data = food_access_DC) |>
  tidy()
```

```{r}
#| label: pulling R Squared

food_access_ND <- food_access |>
  filter(State == "North Dakota")

model_ND <- linear_reg() |>
  set_engine("lm") |>
  fit(`Low Access Numbers.Low Income People.1 Mile` ~ `Vehicle Access.1 Mile` + `Low Access Numbers.Seniors.1 Mile`, data = food_access_ND) 

model_ND |>
  glance() |>
  pull(adj.r.squared)
```

REALLY GOOD MODEL, has high adjusted r squared value,

use

```{r}
#| label: District of Columbia model

food_access$DC <- ifelse(food_access$State == "District of Columbia", 1, 0)

food_access2 <- food_access %>% select(-c(County, State))

model <- glm(DC ~ ., data = food_access2, family = "binomial")


#prob <- predict(model, type = "response")
#roc_obj <- roc(food_access2
               #$DC, prob)
#auc(roc_obj)
```

Hypothesis: We hypothesize that compared to the Northeast, the Midwest will be mostly impacted by having a greater proportion of low income population count beyond 10 miles from supermarket. Since this area is the most rural, there isn't reliable public transportation compared to urban areas to get to supermarkets that are far away.

$H_O$ = The true mean proportional difference between the Midwest and low income population count beyond 10 miles from the supermarket is equal to 0.

$H_a$ = The Midwest has a greater true mean proportion of low income population count beyond 10 miles from the supermarket.


```{r}
#| label: mean for Midwest and Northeast

food_access_reg |>
  filter (region == "Midwest") |>
  summarize (mean = mean(proportion))

food_access_reg |>
  filter (region == "Northeast") |>
  summarize (mean = mean(proportion))

xbar <- 0.1124311 - 0.01673986

xbar
```

The sample mean proportion of low income population count beyond 10 miles from the supermarket is .11 for the Midwest region and .017 for the Northeast region. 

$\bar{x_Midwest} = .11$ 
$\bar {x_Northeast} = .017$ 


The true mean proportion of low income population count beyond 10 miles from the supermarket is .11 for the Midwest region and .017 for the Northeast region. $\bar{x_Midwest} = 0.1124311$ $\bar {x_Northeast} = .017$


```{r}
#| label: visualiation

null_dist <- food_access_reg |>
  filter (region %in% c("Midwest", "Northeast")) |>
  specify (response = proportion, explanatory = region) |>
  hypothesize (null = "independence") |>
  generate (reps = 5000, type = "permute") |>
  calculate (stat = "diff in means", order = c("Midwest", "Northeast"))

```

```{r}

null_dist |>
  ggplot(
    aes(x = stat)) + 
  geom_histogram() +
  geom_vline(xintercept = 0.09569124)

```

```{r}

null_dist |>
  get_p_value(obs_stat = xbar, direction = "greater")

```

Conclusion: The probabiltiy of observing a difference of 0.09569124 between the mean proportion of the Midwest versus the Northeast and low income population count beyond 10 miles from the supermarket is equal to 0, given that the difference is equal to 0, is p = \~0. Because p = \~ 0 \< $alpha$ = .05, we reject the null hypothesis that the true mean proportional difference between the Midwest and low income population count beyond 10 miles from the supermarket is equal to 0. We have strong evidence to conclude the alternate.

**Confidence Interval: True mean proportion for Midwest**

```{r}

set.seed(12345)

boot_df <- food_access_reg |>
  filter (region == "Midwest") |>
  specify(response = proportion) |>
  generate(reps = 5000, type = "bootstrap") |>
  calculate(stat = "mean")

glimpse(boot_df)

```

```{r}

set.seed(12345)
boot_df |>
  summarize(lower = quantile(stat, 0.025),
            upper = quantile(stat, 0.975))

```

```{r}

boot_df |>
  ggplot(
    aes(x = stat)) + 
  geom_histogram() + 
  geom_vline(xintercept = c(0.102609 , 0.1226445)) + 
  labs(title = "95% Confidence Interval",
       subtitle = "True mean price per guest")

```

Interpretation: We are 95% confident that the true mean proportion of low income individuals that live beyond 10 miles from a supermarket in the Midwest is between 0.102609 and 0.1226445.

**Confidence Interval: True mean proportion for Northeast**

```{r}

set.seed(12345)

boot_df <- food_access_reg |>
  filter (region == "Northeast") |>
  specify(response = proportion) |>
  generate(reps = 5000, type = "bootstrap") |>
  calculate(stat = "mean")

glimpse(boot_df)

```

```{r}

set.seed(12345)
boot_df |>
  summarize(lower = quantile(stat, 0.025),
            upper = quantile(stat, 0.975))

```

```{r}

boot_df |>
  ggplot(
    aes(x = stat)) + 
  geom_histogram() + 
  geom_vline(xintercept = c(0.102609 , 0.1226445)) + 
  labs(title = "95% Confidence Interval",
       subtitle = "True mean price per guest")

```

Interpretation: We are 95% confident that the true mean proportion of low income individuals that live beyond 10 miles from a supermarket in the Northeast is between 0.01155406 and 0.02294395.

```{r}
#| label: relationship between low income and vehicle access in the Midwest & Northeast

food_access_reg |>
  filter (region %in% c ("Midwest", "Northeast")) |>
  ggplot (aes (x = proportion, y = proportion_vehicle)) +
  geom_point() + 
  facet_wrap ("region", ncol = 1)
```

There is a positive relationship between the proportion of low income who are 10 miles away from a supermarket and don't have a car who are 10 miles away from a supermarket. As the proportion of low income who are 10 miles away increases so does the proportion for individuals who don't have a car and are 10 miles away.


There is a positive relationship between the proportion of low income who are 10 miles away from a supermarket and don't have a car who are 10 miles away from a supermarket. As the proportion of low income who are 10 miles away increases so does the proportion for individuals who don't have a car and are 10 miles away. Although this graph's proportion for low income and 10 miles away and proportion for no vehicle and 10 miles away is lower. The max it reaches is .4 and .03. While the midwest reaches 1 and .08 for the max. 

--> Sathvika add to results: These two graphs support our claim that Midwest is more impacted as the proportions for the Midwest reach higher numbers for the proportion of low income who are 10 miles away. Moreover, it shows that having no vehicle access contributes to this problem, supporting our claim that Midwest is more impacted by not have reliable access 
to transportation compared to more urban areas like the Northeast. 

```{r}
#| label: relationship between low income and seniors in the Midwest & Northeast

food_access_reg |>
  filter (region %in% c ("Midwest", "Northeast")) |>
  ggplot (aes (x = proportion, y = proportion_seniors)) +
  geom_point() + 
  facet_wrap ("region", ncol = 1)

```

```{r}
#| label: relationship between low income and group quarters in the Midwest & Northeast

food_access_reg |>
  filter (region %in% c ("Midwest", "Northeast")) |>
  ggplot (aes (x = proportion, y = proportion_group)) +
  geom_point() + 
  facet_wrap ("region", ncol = 1)
```






