---
title: "Project title"
subtitle: "Report"
format: html
editor: visual
execute:
  echo: true
---

```{r}
library (tidyverse)
UNICEF_Food <- read.csv("data/UNICEF_Expanded_Global_Databases_FINAL.csv")
slice_head (UNICEF_Food)
```

**Introduction and data**

The source of the dataset is from the United Nations Children's Fund, Division of Data, Analysis, Planning and Monitoring (2022). The dataset includes global statistics on the percentage of children 6-23 months of age who did not consume any fruits and vegetables over the previous day. This dataset originates from October, 2022.

The issue of insufficient fruits and vegetables available in the world traces back decades. According to "Insufficient Consumption of Fruits and Vegetables among Individuals 15 Years and Older in 28 Low- and Middle-Income Countries: What Can Be Done?" Karen S Siegal, expresses that only roughly 20% of those in low and middle income countries are meeting the daily fruit and vegetable average. We want to explore this topic further given how prevalent an issue it is in the world.

Each observation represents three-digit alphabetical codes ISO assigned by the International Organization for Standardization (columns also detail countries, UNICEF regions, UN regions, world bank income group, data source years, and the way that the survey was collected). Previous studies have shown that children from low and low-middle countries had the lowest consumption of fruits and vegetables. Globally, zero vegetable or fruit consumption is very high with 46% of children in the dataset consuming no vegetables or fruit the day before the survey. We will be exploring this concept more with the UNICEF data.

The observations come from qualitative, survey samples from 2015 to 2021.

**Methodology**

Collected through surveys (national/nutrition monitoring and surveillance systems/multiple indicator cluster survey) from 2015 to 2021 and countries with data from this period are included in calculation of regional and global averages with the exception of China where the latest data are from 2013.

We will be using r functions to create graphs to analyze the data and find linear regression for different factors like using parent income, parent education, and wealth quantile.

Using this data, we hope to develop an inference to predict if lower or higher incomes have lower or higher consumption of fruits and vegetables for children.

**Results**

Are there greater gender differences in the consumption of zero vegetable or fruit for low income countries than in higher income countries?

We hypothesize that there are male infants are better fed than female infants in lower-income countries whereas this gender gap is less prevalent in upper-middle income countries. The variables involved in our research question are numerical, with the indicator being the percentage of children 6-23 months of age who did not consume any vegetables or fruits during the previous day.

We hypothesize that there are male infants are better fed than female infants in lower-income countries whereas this gender gap is less prevalent in upper-middle income countries. The variables involved in our research question are numerical, with the indicator being the percentage of children 6-23 months of age who did not consume any vegetables or fruits during the previous day.

```{r}
UNICEF_Food |>
  group_by(WB_Latest) |>
  summarise(mean_male = mean(male_r, na.rm = T), 
            mean_female = mean(female_r, na.rm = T))
```

There seems to be a significant difference between female and male consumption of fruit based on the previous day depending on the income categorization of their country. Around 14% of males in the high income category did not consume fruit in the previous day. Around 24% of females in the high income category did not consume fruit in the previous day.

```{r}
UNICEF_Food |>
  filter(WB_Latest == "Low Income") |>
  summarise(mean_urban = mean(urban_r, na.rm = T), 
            mean_rural = mean(rural_r, na.rm = T))
```

```{r}
UNICEF_Food |>
  filter(WB_Latest == "Low Income") |>
  summarise(mean_6to11 = mean(month_6_11_r),
            mean_12to15 = mean(month_12_15_r), 
            mean_16to19 = mean(month_16_19_r))
```

```{r}
UNICEF_Food |>
  filter(WB_Latest == "High Income") |>
  summarise(mean_6to11 = mean(month_6_11_r, na.rm = T),
            mean_12to15 = mean(month_12_15_r, na.rm = T), 
            mean_16to19 = mean(month_16_19_r, na.rm = T))
```

In high income countries, age does not seem to play as an important factor in consumption of fruit.

```{r}
UNICEF_Food |>
  ggplot(aes(x = UNRegion, fill = WB_Latest, na.rm = T)) + 
  geom_bar() 
```
